Object-based Auditory Scenes Similarity Retrieval and Classification With Wavelet Scattering

### 2016-07-12 paper/structureScene.tex
Edit motivations

As stated in the introduction, the field of ecoacoustics is relatively young, but significant works are worth mentioning.
One avenue of research, originated in bioacoustics, intends to retrieve the species that produced a given sonic event of relatively short duration, in order to link animal vocalization with behavior\cite{au2008principles, stowell2014large}.
A more recent paradigm processes the audio stream at hand in a holistic way, as it describes the auditory environment over larger temporal scales, without assuming that one single species is present throughout the recording. Automated systems belonging to this paradigm aim at inferring global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010}, migration patterns \cite{Obrist2010}, as well as regions of particular interest before human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating}.

The same dichotomy applies to the field of auditory scene analysis, wherein the interest may either be 1) to isolate and identify precise events, \ie acoustic event detection (AED), or 2) to predict global attributes assigned to large time spans of audio, \ie acoustic scene classification (ASC).
In the latter, most state-of-the-art systems roughly follow the bag of frame approach (BOF) initiated in \cite{aucouturier2007bag}, which consists in modeling audio by high-level summary statistics computed from a set of local features.
As it enjoys a high genericity, this approach is found in many audio classification tasks besides ASC.
Yet, the specific implementation with gaussian mixture models (GMMs) of Mel-frequency cepstral coefficients (MFCCs) it is demonstrated to be perform equivalently compared to a mere averaging of the features  \cite{lagrange:hal-01082501}.
Indeed, the typical morphology of acoustic scenes is of a "skeleton of events on a bed of textures" \cite{nelken_ear_2013}, \ie few impulsive sounds superimposed upon a stationary background.
Therefore, it is arguably more beneficial to focus on the temporal evolution of auditory scenes rather on the joint probability distribution of short-term features.

Interestingly, this statement is validated by auditory psychology as well as sound synthesis based  on summary statistics \cite{mcdermott2013summary}. Indeed studies in cognitive psychology investigating urban sound environment quality pointed out that low level features such as (perceived or measured) global sound level, are not sufficient to fully characterized an acoustic scene \cite{guyot2005urban,kang2006urban}. Instead several papers have reported that some cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003}  relies upon high level cognitive attributes such as the identification and assessment of the sound source which compose the scene. It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high level cognitive classes, \textit{i.e.} the presence of birds are strong pleasantness indicators and very likely to be heard in parks in urban areas \cite{lafay:hal-01111782}. As a consequence, research in sound perception now focus on the study of the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}.

Although the precise set of events occuring within a given auditory stream may not be discernable, even for trained humans, research shows that the focus on a few events, so called markers,  sufficed to reliably predict many high-level attributes \cite{lafayPartI}.
For example, the presence of a few bird calls in a smooth traffic hub is likely to be tagged by humans as a park scene in urban areas.

>>>>

One avenue of ecoacoustics research, originating in bioacoustics, attempts to identify the species that produced a specific sonic event of short duration in order to link animal vocalization with behavior\cite{au2008principles, stowell2014large} \ja{Is this the only reason?}.
A more recent paradigm processes the audio stream in a holistic way over larger time scales, without assuming that a single species is present throughout the recording. Automated systems belonging to this paradigm aim to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010}, migration patterns \cite{Obrist2010}, as well as regions of particular interest before human inspection \cite{rosenstock2002landbird,diefenbach2007incorporating} \ja{What does this mean?}.

The same division is found in the field of auditory scene analysis, wherein the interest may either be 1) to isolate and identify precise events, \ie acoustic event detection (AED), or 2) to predict global attributes assigned to large time spans of audio, \ie acoustic scene classification (ASC).
In the latter \ja{What about former? Why are we discussing both if we are going to focus on one? Shouldn't we just say that at some point?}, most state-of-the-art systems more or less follow the bag-of-frames (BOF) approach introduced by Acouturier et al. \cite{aucouturier2007bag}, where a scene is modeled by high-level summary statistics computed from local features.
Being a very generic method, this approach has been applied to many audio classification tasks besides ASC.
Yet the proposed implementation using Gaussian mixture models (GMMs) of mel-frequency cepstral coefficients (MFCCs) has been demonstrated comparably to a mere averaging of the features \cite{lagrange:hal-01082501}.
Indeed, the typical morphology of acoustic scenes is of a ``skeleton of events on a bed of textures'' \cite{nelken_ear_2013}, with a few discrete sound events superimposed upon a stationary acoustic background.
Therefore, it is arguably more beneficial to focus on the large-scale temporal evolution of auditory scenes rather on the joint probability distribution of short-term features.

This statement has some support in auditory psychology as well as sound synthesis based on summary statistics \cite{mcdermott2013summary}. Indeed, studies in cognitive psychology investigating urban sound environment quality pointed out that low-level features such as (perceived or measured) global sound level \ja{Other things like spectral envelope too?}, are not sufficient to fully characterize an acoustic scene \cite{guyot2005urban,kang2006urban}. Rather, it seems that cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003} rely upon higher-level cognitive attributes such as the identities of sound sources which compose the scene. \ja{This argument is semantically a little tricky, since technically the low-level features already contain all this high-level information, it's just that the simple models we use to describe them fail to capture this high-level structure} It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high-level cognitive classes, \textit{e.g.} the presence of birds are very likely to be heard in parks in urban areas and so are strong pleasantness indicators \cite{lafay:hal-01111782}. Consequently, research in sound perception is now focused the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}.

Although the complete set of events occuring within a given auditory stream may not be discernable, even to human expects, research has shown that a small set of events, so-called markers, suffice to reliably predict many high-level attributes \cite{lafayPartI}.
For example, the presence of a few bird calls in a smooth traffic hub is likely to be tagged by humans as a park scene in urban areas. \ja{This is very similar to the last few sentences of the last paragraph. What are we trying to say differently here?}
